A Data-Driven Approach to Power Converter Control via Convex Optimization

### 2017-02-28 Papers/Loop Shaping Voltage Loop (CERN)/Latex/Draft_2/Draft_2.tex
Update Draft_2.tex

\author{Achille Nicoletti$^{1}$ and Michele Martino$^{1}$ and Alireza Karimi$^{2}$% <-this % stops a space
\thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$Achille Nicoletti and Michele Martino are both with the European Organization for Nuclear Research (CERN), 
 CH-1211 Geneva 23, Switzerland
        {\tt\small  e-mail: achille.nicoletti@cern.ch / michele.martino@cern.ch}}%
\thanks{$^{2}$Alireza Karimi is with Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Institute of Mechanical Engineering (IGM),
   CH-1015 Lausanne, Switzerland
        {\tt\small e-mail: alireza.karimi@epfl.ch}}%
}

@@

The method proposed in this paper is an extension of \cite{KK16} to the model reference control problem and invokes a new data-driven control design scheme for specific power converter applications at CERN. Currently, the load (i.e., the magnet of the particle accelerator) is approximated as a simple first-order system (series $RL$ circuit); however, it has been shown in \cite{KZ98}, \cite{NW95} and \cite{NS79} that the vacuum chamber of the magnet is also frequency dependent (mainly due to the coupling effects of the beam and deformations in the chamber geometry). In fact, the authors in \cite{NW95} assert that even small deformations in the beam pipe can drastically change the impedance of the system and create stability issues. Therefore, it is appropriate to consider a data-driven based design to regulate the power converter control system.  In this work, it is shown that a model reference design in the $\mathcal{H}_\infty$ or $\mathcal{H}_2$ sense can be formulated through a convex optimization problem; however, the methods described in this paper can also be applied to mixed $\mathcal{H}_2$ and $\mathcal{H}_\infty$ problems (i.e., minimizing the norm of weighted sensitivity functions). Additionally, with certain trivial conditions on the controller parameters, it is shown that the closed-loop stability is guaranteed.

@@

\section{System Framework}
\label{sec:system}
\begin{figure*}
\centering
\includegraphics[width=1.7\columnwidth]{../pics/voltage_loop_schematic_color_coded}
\caption{Power converter control system schematic. The magnet represents the load of the process while the DSP implements the control strategy. The blue portion represents the power electronics of the converter; the yellow portion represents the analog process of the converter with the load; the red portion represents the digital-signal processor (DSP) which implements the control strategy; and the green block represents the measurement board which reads the desired control variables and sends the information to the DSP for processing.}
\label{fig:circuit}
\end{figure*}
A schematic of the power converter control system is shown in Fig.~\ref{fig:circuit}. Specific portions of the schematic are color-coded in order to construct the associated block diagram. Note that there are two power converter branches assembled in series in this schematic; however, both branches possess the same components with identical values. Therefore, the analysis and control strategy can be performed in one branch and then scale the controller parameters accordingly (i.e., by a factor of 2). The input voltage for one of the high frequency (HF) filters will be denoted as $v_i$ while the output voltage of the filter (for one branch) will be denoted as $v_0$. The TF from $v_i$ to $v_0$ describes the output filter of the power converter for one branch. For general applications at CERN, the objective is to control the current in the magnet such that the error between this output current and a delayed version of the reference current is minimized. However, the objective of this paper is to design and implement a controller to shape the FRF between the reference voltage and $v_0$. This intermediate control is performed due to the fact that the output typically possesses large resonances, and a controller is required to attenuate these resonances in order to ensure better performance in the current loop. These large resonances are present because the values of the filter resistances are small (in order to reduce the power losses). Large resistance values can attenuate the resonances at the price of larger power losses. Another reason for implementing this type of control structure at CERN is for commissioning purposes; if/when faults occur in the system, each loop can be verified one at a time to ensure quality operation.

The block diagram of the control strategy is shown in Fig.~\ref{fig:damping_loop}. This block diagram is color-coded to correlate with the structure of the schematic in Fig.~\ref{fig:circuit}. The analog-to-digital converter (ADC) is used to process the information from the analog system to the DSP; $G_0(s)$ represents the TF from $v_i$ to $v_0$; $H(s)$ represents the TF from $v_0$ to $i_c$ (the current through the capacitor in the HF filter); $P_1(s)$ and $P_2(s)$ are sensors (i.e., low-pass filters); $k_i$ and $k_u$ are the control variable gains; $\tau \: [s]$ is an actuation delay that is used to approximate the dynamics of the power electronics; and $C(s)$ is a PI controller, i.e., $$C(z^{-1}) = k_{p_1} + k_{p_2}\frac{t_s z^{-1}}{1-z^{-1}}$$ where $t_s \: [s]$ is the sampling rate of the DSP. The ADCs have very fast dynamics relative to the other subsystems in the loop; for this reason, it is appropriate to consider the ADCs as pure time delays (which will be denoted with a delay $\tau_{a} \: [s]$). The transfer function analyzer (TFA) is used for identification purposes, and will be discussed in the case study. As stated in the Remark in the previous section, the FRFs of continuous-time and discrete-time systems are nearly identical when the Nyquist-Shannon sampling theorem is met. Since the proposed method invokes a data-driven method using the FRF of a system, all functions discussed hereafter will be FRFs.

@@

In order to simplify the system framework, let us define the following quantities:
\begin{equation} \label{eq:new_var}
\begin{aligned}
k_{m_1} &:= k_{p_1}(1+k_u) ;  &k_{m_2} &:= k_{p_2}(1+k_u)  \\
H_1(k) &:= k_iP_1H;  &H_2(k) &:= k_uP_2e^{-j\omega \tau_{a}} \\
H_3(k) &:=  P_2C_m(k)e^{-j\omega\tau_{a}}; &G &:= G_0e^{-j\omega\tau} \\ 
\end{aligned}
\end{equation}
\begin{equation*}
C_m(k) := k_{m_1}+k_{m_2}\frac{t_se^{-j\omega}}{1-e^{-j\omega}}
\end{equation*}
where $k = [k_i,k_u,k_{m_1},k_{m_2}]$. Let the FRF from $v_d$ to $v_0$ be denoted as $T^{\prime}(k)$ and the FRF from $v_r$ to $v_0$ be denoted as $T(k)$. These closed-loop FRFs can easily be derived as follows:

@@

\begin{table}[]
\centering
\caption{System characteristics of SATURN power converter}
\label{tab:specs}
\begin{tabular}{c|c|c}
                                                                               & Value                      & Units \\ \hline
\begin{tabular}[c]{@{}c@{}}$U_0$\\ (Output Voltage)\end{tabular}               & $700$                      & V     \\ \hline
\begin{tabular}[c]{@{}c@{}}$I_0$\\ (Output Current)\end{tabular}               & $850$                      & A     \\ \hline
Input Power                                                                    & $18$                       & kV    \\ \hline
Input Frequency                                                                & $50$                       & Hz    \\ \hline
\begin{tabular}[c]{@{}c@{}}IGBT equivalent \\ switching frequency\end{tabular} & $5000$                     & Hz    \\ \hline
Transformer Weight                                                             & $4500$                     & kg    \\ \hline
DSP sampling rate                                                             & $5000$                     & Hz    \\ \hline
Control Type                                                                   & RegFGC3 Type 10 / Ethernet \cite{CKS11} & $-$  
\end{tabular}
\end{table}

@@

\subsection{Controller Synthesis}
For this application, it was desired to shape $T^{\prime}(k)$ and $T(k)$ such that a bandwidth of 300 Hz and 500 Hz is attained, respectively (where both achieve a damping of $\zeta = 0.8$). Note that in both of these TF's, a pure delay appears in the numerator; therefore, the performance of the loop-shape algorithm can be improved by specifying a desired loop-shape $T_d(s) = T_d^*(s) e^{-s\tau}$ where $T_d^*(s)$ was selected as a standard second order process, i.e.,
\begin{equation}
T_d^*(s) = \frac{\omega_d^2}{s^2 + 2\zeta \omega_d s + \omega_d^2}
\end{equation}
where $\zeta$ is the damping factor and
\begin{equation*}
\omega_d = \frac{2 \pi f_d}{\sqrt{1-2\zeta^2 + \sqrt{2-4\zeta^2 + 4\zeta^4}}}
\end{equation*}
where $f_d$ is the desired closed-loop bandwidth. In other words, $T_d$ is simply a delayed version of the second order process. There is some uncertainty associated with the delay $\tau$; the range of this delay is $\tau \in [0,200]\mu s$ with a nominal value of $100 \mu s$. Therefore, for tracking purposes, $\tau = 100 \mu s$ in $T_d(s)$ can be selected as a viable performance criteria. Note that selecting $\tau$ to be another value within the uncertain range will not significantly impact the response. The critical values in the desired response are $f_d$ and $\zeta$.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../pics/d_loop_mag.eps}
\caption{The closed-loop response of $|T^{\prime}|$; data-driven method with $\mathcal{H}_{\infty}$ performance (red line); data-driven method with $\mathcal{H}_2$ performance (blue line); model-based approach with $\mathcal{H}_2 / \mathcal{H}_\infty$ performance (yellow line); -3 dB point representing 300 Hz bandwidth (dashed-black line).}
\label{fig:d_loop}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../pics/v_loop_mag.eps}
\caption{The closed-loop response of $|T|$; data-driven method with $\mathcal{H}_{\infty}$ performance (red line); data-driven method with $\mathcal{H}_2$ performance (blue line); model-based approach with $\mathcal{H}_2 / \mathcal{H}_\infty$ performance (yellow line); -3 dB point representing $500$ Hz bandwidth (dashed-black line).}
\label{fig:v_loop}
\end{figure}

>>>>

\author{Achille Nicoletti$^{1,2}$ and Michele Martino$^{1}$ and Alireza Karimi$^{2}$% <-this % stops a space
\thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$Achille Nicoletti and Michele Martino are both with the European Organization for Nuclear Research (CERN), 
 CH-1211 Geneva 23, Switzerland
        {\tt\small  e-mail: achille.nicoletti@cern.ch / michele.martino@cern.ch}}%
\thanks{$^{2}$Alireza Karimi is with Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Institute of Mechanical Engineering (IGM),
   CH-1015 Lausanne, Switzerland
        {\tt\small e-mail: alireza.karimi@epfl.ch}}%
}

@@

The method proposed in this paper is an extension of \cite{KK16} to the model reference control problem and invokes a new data-driven control design scheme for specific power converter applications at CERN. Currently, the load (i.e., the magnet of the particle accelerator) is often approximated as a simple first-order system (series $RL$ circuit) even though it has been shown in \cite{Shafer1}, \cite{Shafer2} and \cite{Ravaioli1} that both real and imaginary part of its impedance are ferquency dependent and its behavior is also affected by the vacuum chamber of the magnet. Therefore, it is appropriate to consider a data-driven based design to regulate the power converter control system.  In this work, it is shown that a model reference design in the $\mathcal{H}_\infty$ or $\mathcal{H}_2$ sense can be formulated through a convex optimization problem; however, the methods described in this paper can also be applied to mixed $\mathcal{H}_2$ and $\mathcal{H}_\infty$ problems (i.e., minimizing the norm of weighted sensitivity functions). Additionally, with certain trivial conditions on the controller parameters, it is shown that the closed-loop stability is guaranteed.

@@

\section{System Framework}
\label{sec:system}
\begin{figure*}
\centering
\includegraphics[width=1.7\columnwidth]{../pics/voltage_loop_schematic_color_coded}
\caption{Power converter control system schematic. The magnet represents the load of the process while the DSP implements the control strategy. The blue portion represents the power electronics of the converter; the yellow portion represents the analog process of the converter with the load; the red portion represents the digital-signal processor (DSP) which implements the control strategy; and the green block represents the measurement board which reads the desired control variables and sends the information to the DSP for processing.}
\label{fig:circuit}
\end{figure*}
A schematic of the power converter with its control system is shown in Fig.~\ref{fig:circuit}. Specific portions of the schematic are color-coded in order to construct the associated block diagram. Note that there are two power converter branches assembled in series in this schematic; however, both branches possess the same components with identical values. Therefore, the analysis and control strategy can be performed in one branch and then scale the controller parameters accordingly (i.e., by a factor of 2). The input voltage for one of the high frequency (HF) filters will be denoted as $v_i$ while the output voltage of the filter (for one branch) will be denoted as $v_0$. The TF from $v_i$ to $v_0$ describes the output filter of the power converter for one branch; the HF filter is needed to remove the high frequency components of the PWM voltage at the output of the SMPS power stage (SMPS Switched Mode Power Supply). For general applications at CERN, the objective is to control the current in the magnet such that the error between this output current and a delayed version of the reference current is minimized. However, the objective of this paper is to design and implement a controller to shape the FRF between the reference voltage and $v_0$. This intermediate control is performed due to the fact that the HF filter typically possesses large resonances, and a controller is required to attenuate these resonances in order to ensure better performance in the current loop. Large resonances are present because the values of the filter resistances are kept small in order to reduce the power losses. Large resistance values can attenuate the resonances at the price of larger power losses and would therefore compromise energy efficiency. Another reason for implementing this type of control structure at CERN is for commissioning purposes; if/when faults occur in the system, each loop can be verified one at a time to ensure quality operation.

The block diagram of the control strategy is shown in Fig.~\ref{fig:damping_loop}. This block diagram is color-coded to correlate with the structure of the schematic in Fig.~\ref{fig:circuit}. The analog-to-digital converter (ADC) is used to process the information from the analog system to the DSP; $G_0(s)$ represents the TF from $v_i$ to $v_0$; $H(s)$ represents the TF from $v_0$ to $i_c$ (the current through the capacitor in the HF filter); $P_1(s)$ and $P_2(s)$ are sensors (i.e., low-pass filters); $k_i$ and $k_u$ are the control variable gains; $\tau \: [s]$ is an actuation delay that is used to approximate the dynamics of the power electronics; and $C(z^{-1})$ is a PI controller, i.e., $$C(z^{-1}) = k_{p_1} + k_{p_2}\frac{t_s z^{-1}}{1-z^{-1}}$$ where $t_s \: [s]$ is the sampling rate of the DSP. The ADCs have very fast dynamics relative to the other subsystems in the loop; for this reason, it is appropriate to consider the ADCs as pure time delays (which will be denoted with a delay $\tau_{a} \: [s]$). The transfer function analyzer (TFA) is used for identification purposes, and will be discussed in the case study. As stated in the Remark in the previous section, the FRFs of continuous-time and discrete-time systems are nearly identical when the Nyquist-Shannon sampling theorem is met. Since the proposed method invokes a data-driven method using the FRF of a system, all functions discussed hereafter will be FRFs.

@@

In order to simplify the system framework, let us define the following quantities:
\begin{equation} \label{eq:new_var}
\begin{aligned}
k_{m_1} &:= k_{p_1}(1+k_u) ;  &k_{m_2} &:= k_{p_2}(1+k_u)  \\
H_1(k) &:= k_iP_1H;  &H_2(k) &:= k_uP_2e^{-j\omega \tau_{a}} \\
H_3(k) &:=  P_2C_m(k)e^{-j\omega\tau_{a}}; &G &:= G_0e^{-j\omega\tau} \\ 
\end{aligned}
\end{equation}
\begin{equation*}
C_m(k) := k_{m_1}+k_{m_2}\frac{t_se^{-j\omega t_{s}}}{1-e^{-j\omega t_{s}}}
\end{equation*}
where $k = [k_i,k_u,k_{m_1},k_{m_2}]$. Let the FRF from $v_d$ to $v_0$ be denoted as $T^{\prime}(k)$ and the FRF from $v_r$ to $v_0$ be denoted as $T(k)$. These closed-loop FRFs can easily be derived as follows:

@@

\begin{table}[]
\centering
\caption{System characteristics of SATURN power converter}
\label{tab:specs}
\begin{tabular}{c|c|c}
                                                                               & Value                      & Units \\ \hline
\begin{tabular}[c]{@{}c@{}}$U_0$\\ (Output Voltage)\end{tabular}               & $700$                      & V     \\ \hline
\begin{tabular}[c]{@{}c@{}}$I_0$\\ (Output Current)\end{tabular}               & $850$                      & A     \\ \hline
AC Input                                                                       & $18$                       & kV    \\ \hline
Input Frequency                                                                & $50$                       & Hz    \\ \hline
\begin{tabular}[c]{@{}c@{}}IGBT equivalent \\ switching frequency\end{tabular} & $5000$                     & Hz    \\ \hline
DSP sampling rate                                                              & $5000$                     & Hz    \\ \hline
Control Type                                                                   & RegFGC3 Type 10 / Ethernet \cite{CKS11} & $-$  
\end{tabular}
\end{table}

@@

\subsection{Controller Synthesis}
For this application, it was desired to shape $T^{\prime}(k)$ and $T(k)$ such that a bandwidth of 300 Hz and 500 Hz is attained, respectively (where both achieve a damping of $\zeta = 0.8$). Note that in both of these TF's, a pure delay appears in the numerator; therefore, the performance of the loop-shape algorithm can be improved by specifying a desired loop-shape $T_d(s) = T_d^*(s) e^{-s\tau}$ where $T_d^*(s)$ was selected as a standard second order process, i.e.,
\begin{equation}
T_d^*(s) = \frac{\omega_d^2}{s^2 + 2\zeta \omega_d s + \omega_d^2}
\end{equation}
where $\zeta$ is the damping factor and
\begin{equation*}
\omega_d = \frac{2 \pi f_d}{\sqrt{1-2\zeta^2 + \sqrt{2-4\zeta^2 + 4\zeta^4}}}
\end{equation*}
where $f_d$ is the desired closed-loop bandwidth. There is some uncertainty associated with the delay $\tau$; the range of this delay is $\tau \in [0,200]\mu s$ with a nominal value of $100 \mu s$. Therefore, for tracking purposes, $\tau = 100 \mu s$ in $T_d(s)$ can be selected as a viable performance criteria. Note that selecting $\tau$ to be another value within the uncertain range will not significantly impact the response. The critical values in the desired response are $f_d$ and $\zeta$.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../pics/d_loop_mag.eps}
\caption{Measured closed-loop response of $|T^{\prime}|$; data-driven method with $\mathcal{H}_{\infty}$ performance (red line); data-driven method with $\mathcal{H}_2$ performance (blue line); model-based approach with $\mathcal{H}_2 / \mathcal{H}_\infty$ performance (yellow line); -3 dB point representing 300 Hz bandwidth (dashed-black line).}
\label{fig:d_loop}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../pics/v_loop_mag.eps}
\caption{Measured closed-loop response of $|T|$; data-driven method with $\mathcal{H}_{\infty}$ performance (red line); data-driven method with $\mathcal{H}_2$ performance (blue line); model-based approach with $\mathcal{H}_2 / \mathcal{H}_\infty$ performance (yellow line); -3 dB point representing $500$ Hz bandwidth (dashed-black line).}
\label{fig:v_loop}
\end{figure}
